---
title: "Food Insecurity Index Creation"
author: "Chungmann Kim"
date: "`r Sys.Date()`"
categories: ["R"]
tags: ["LSMS", "R Markdown", "Food Security", "Malawi"]
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,dpi=300)
```

```{r, include=FALSE}
pack <- c('tidyverse','haven','janitor','zoo')
install.packages(setdiff(pack, rownames(installed.packages())))
sapply(pack,require,character.only=TRUE)

```

# Wrangling Living Standards Measurement Survey(LSMS) Workbank Data 

This r code chunk links different data sources together and creates a single dataframe with multiple food insecurity measures(e.g. Food consumption score(FCS), Household Diet Diverssity Score(HDDS), reduced Coping Strategy Index (rCSI)) computed and several other rosters that identify each observation.

```{r chunk1, warning = FALSE, echo=TRUE}
#####################################################################
########################    Malawi 2010    ##########################
#####################################################################

######################################################################
##################### 1. Food security index #########################
######################################################################

### (1) Find data and recode

path = "data/Malawi_2010/"

mw10.food <- read_dta(file = paste(path,"Household_Batch3of5_DTA/HH_MOD_G2.dta",sep = "" ) )

mw10.food.category.days = mw10.food %>% 
  # Recoding of outliers (change 8 days to 7 days)
  mutate( hh_g08c  = if_else(hh_g08c>7,7,hh_g08c) ) %>%
  #Combining Cereals and roots (Category A and Category B)
  mutate(hh_g08a = if_else(hh_g08a=="A" | hh_g08a=="B","AB",hh_g08a) ) %>%
  # combine A and B 
  group_by(case_id,ea_id,hh_g08a) %>%
  summarise( hh_g08c =max (hh_g08c))


###(2) compute FCS

######################################################################
#NOTES ON THE WEIGHTS OF THE DIFFERENT FOOD CATEGORIES
# A Cereals, Grains and Cereal Products: Weight = 2
# B Root, Tubers and Plantains: Weight = 2
# C Nuts and Pulses: Weight = 3
# D Vegetables: Weight = 1
# E Meat, Fish and Animal Products: Weight = 4
# F Fruits => weight = 1
# G Milk/Milk Products: Weight = 4
# H Fats/Oil => Weight = 0.5
# I Sugar/Sugar Products/Honey: Weight = 0.5
# J Spices/Condiments: Weight = 0
######################################################################
FWeight.mapping= read.table(text = "
category FWeight 
                            AB 2
                            C 3
                            D 1
                            F 1
                            E 4
                            G 4
                            H 0.5
                            I 0.5
                            J 0", header = TRUE)



FCS = mw10.food.category.days %>%
  mutate(
    # find the weight from 
    FWeight = with(FWeight.mapping,FWeight[hh_g08a %in% category])
  ) %>%
  mutate( FCS = hh_g08c * FWeight ) %>% 
  group_by(case_id,ea_id) %>%
  summarise( FCS  = sum (FCS,na.rm = TRUE) )

FCS = FCS %>% filter(FCS!=0)
mw10.merged = FCS


### (3) recodeand compute HDDS 

################################################################################
#A diet diversity score is a household-measure of food security that captures ///
#something about the quality of a diet. It is calculated by counting the number///
#of foods or food groups from which a household acquired food over the survey ///
#reference period (24 hours). 
################################################################################

HDDS = mw10.food.category.days %>% 
  #Exclude SUGAR and SPICES
  filter(hh_g08a!="I" & hh_g08a!="J") %>%
  filter(  !is.na(hh_g08c) ) %>%
  mutate( HDDS = if_else( hh_g08c>=1 ,1,0 )  ) %>%
  group_by(case_id,ea_id) %>%
  summarise(HDDS = sum(HDDS,na.rm = TRUE))

HDDS = HDDS %>% filter(HDDS!=0)

mw10.merged = left_join(mw10.merged,HDDS, by=c("ea_id","case_id"))

### (4) Find, recode, and compute rCSI

################################################################################
#In the past 7 days, did you worry that your household would not have enough food
#hh_h02a: "In the past 7 days, how many days have you or someone in your ///
#household had to: Rely on less preferred and/or less expensive foods?"(WGT1)
#hh_h02b: "In the past 7 days, how many days have you or someone in your ///
#household had to: Limit portion size at mealtimes?" (WGT1)
#hh_h02c: "In the past 7 days, how many days have you or someone in your ///
#household had to: Reduce number of meals eaten in a day?" (WGT2)
# hh_h02d "In the past 7 days, how many days have you or someone in your ///
#household had to: Restrict consumption by adults in order for small ///
#children to eat?" (WGT2)
#hh_h02e "In the past 7 days, how many days have you or someone in your ///
#household had to: Borrow food, or rely on help from a friend or ///
#relative?" (WGT2)
################################################################################

mw10.rcsi <- read_dta(file = paste(path,"Household_Batch3of5_DTA/HH_MOD_H.dta",sep = "" ) )

##Constructing rCSI
rCSI = mw10.rcsi %>%
  mutate(rCSI = 1*hh_h02a + 1*hh_h02b + 2*hh_h02c + 2*hh_h02d +2*hh_h02e ) %>%
  mutate(rCSI = if_else(rCSI>42 , 42 ,rCSI) ) %>% 
  dplyr::select(case_id,ea_id,rCSI)

mw10.merged = left_join(mw10.merged,rCSI, by=c("ea_id","case_id"))


#######################################################################
# 2. PICK UP GEOLOCATION and TIME(Year and month) and hhweight ########
#######################################################################

# pick up urban/rual, hhweight, FS_month, FS_year
mw10.region<- read_dta(file = paste(path,"Household_Batch1of5_DTA/HH_MOD_A_FILT.dta",sep = "" ))

mw10.region.rural = mw10.region %>% 
  mutate(rural = case_when(
    reside == 1 ~ 0,
    reside == 2 ~ 1)
  ) %>% 
  mutate(FS_year = hh_a23c_1) %>%
  mutate(FS_month = hh_a23b_1) %>%
  dplyr::select(case_id,ea_id,rural, FS_year, FS_month, hh_wgt) %>% rename(hhweight=hh_wgt)

mw10.merged = left_join(mw10.merged,mw10.region.rural, by=c("ea_id","case_id"))

#######################################################################
# 3. Pick up geolocation (lat/lon) and create "round"column ###########
#######################################################################

mw10.geo <- read_dta(file = paste(path,"HouseholdGeovariables.dta",sep = "" ))
mw10.geo.clean = mw10.geo %>%
  dplyr::select(case_id,ea_id,lat_modified,lon_modified)

mw10.merged = left_join(mw10.merged,mw10.geo.clean, by=c("ea_id","case_id")) %>%
  mutate(round = 2010) %>%
  rename(HHID = case_id)

  
#######################################################################
# 4. Check output and export ##########################################
#######################################################################
head(mw10.merged)
```

# Create Simple Plot
I create a dummy variables which codify whether a household is Food Insecure or not based on two different meassures - FCS, rCSI using a certain threshold.
It turns out that using FCS generates more food inscure household than rCSI.

```{r plot, warning = FALSE, echo=TRUE}
plotdf = mw10.merged %>%
  mutate(rCSI_FI = if_else(rCSI > 17 , 1 , 0) ) %>% 
  mutate(FCS_FI = if_else(FCS < 35 , 1 ,0) ) %>%
  pivot_longer(
  cols=rCSI_FI:FCS_FI,
  names_to = "FI",
  values_to = "FI_index"
  )

ggplot(plotdf, aes(x= FI_index, color = FI)) +
  geom_bar() + facet_grid(rural~FI) +
  labs(title = "Count FCS based FI index VS rCSI based FI index",
  x = "FCS vs rCSI", y = "Count", color = "Food Insecurity Base")
```
